# Ch. 6: Tasks & Quizzes {.unnumbered}
---

::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q6.1**]{style="color:green;"} What types of data were collected as part of this study?

```{r}
#| echo: false
#| label: "Q6.1"
library(checkdown)
check_question(c("Participants' results on a grammar test", "Participants' results on a vocabulary test", "Participants' results on a collocations test", "Sociodemographic information about the participants including their age and level of education", "Participants' results on a language analytic ability test", "Participants' results on a nonverbal intelligence test"), 
               options = c("Participants' results on a grammar test", "Participants' results on a vocabulary test", "Participants' results on a collocations test", "Sociodemographic information about the participants including their age and level of education", "Participants' results on a language analytic ability test", "Participants' results on a nonverbal intelligence test"), 
               type = "check",
button_label = "Check answer",
random_answer_order = TRUE,
q_id = "Q6.1", 
right = "Well done!",
wrong = "That's a good start, but there are more...")
```

[**Q6.2**]{style="color:green;"} On average, how did the English L2 speakers perform compared to the native speakers?

```{r}
#| echo: false
#| label: "Q6.2"

check_question("On average, native speakers outperformed L2 speakers on all language tasks, with the most significant difference observed in collocation tasks.",
               options = c("On average, L2 speakers performed better on grammar tasks than native speakers", "On average, native speakers outperformed L2 speakers on all language tasks, with the most significant difference observed in collocation tasks.", "On average, both groups performed equally well on grammar and vocabulary tasks.", "The observed differences in performance between native speakers and L2 speakers were not statistically significant."), 
               type = "radio",
button_label = "Check answer",
random_answer_order = TRUE,
q_id = "Q6.2", 
right = "That's right.",
wrong = "No, read the abstract again.")
```

[**Q6.3**]{style="color:green;"} Did all native speakers perform better than the L2 speakers in the English vocabulary and grammar tests?

```{r}
#| echo: false
#| label: "Q6.3"

check_question("No, while L1 speakers generally performed better, some L2 speakers demonstrated equally high proficiency in grammar and vocabulary.",
               options = c("No, while L1 speakers generally performed better, some L2 speakers demonstrated equally high proficiency in grammar and vocabulary.", "Yes, the L1 speakers clearly outperformed the L2 speakers in both grammar and vocabulary.", "This study only looked at average trends, so no conclusive statement can be made about individual participants."), 
               type = "radio",
button_label = "Check answer",
random_answer_order = TRUE,
q_id = "Q6.3", 
right = "That's right.",
wrong = "No, read the abstract again.")
```

:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q6.4**]{style="color:green;"} On which repository/repositories can the data be found?

```{r}
#| echo: false
#| label: "Q6.4"

check_question("IRIS",
               options = c("Zenodo", "database.org", "IRIS", "ResearchGate", "All of the above."), 
               type = "radio",
button_label = "Check answer",
random_answer_order = FALSE,
q_id = "Q6.4", 
right = "That's right!",
wrong = "Are you sure?")
check_hint("Check the links listed in Appendix S4.", 
           hint_title = "🐭 Click on the mouse for a hint.")

```

:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q6.5**]{style="color:green;"} In which data format are these two files saved?

```{r}
#| echo: false
#| label: "Q6.5"

check_question("They are both .csv files.",
               options = c("They are both .csv files.", "They are both .txt files.", "They are both .html files.", "They are both .xslx files."), 
               type = "radio",
button_label = "Check answer",
q_id = "Q6.5", 
random_answer_order = TRUE,
right = "That's right! 🎉",
wrong = "Are you sure? 🤨")

```

:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q6.6**]{style="color:green;"} Which character is used to separate the values in the file `L1_data.csv`?

```{r}
#| echo: false
#| label: "Q6.6"

check_question("comma",
               options = c("comma", "tab", "double quotation mark", "colon", "space"), 
               type = "radio",
button_label = "Check answer",
random_answer_order = TRUE,
q_id = "Q6.6", 
right = "That's right!",
wrong = "Are you sure?")

```

[**Q6.7**]{style="color:green;"} Which character is used to delineate the values?

```{r}
#| echo: false
#| label: "Q6.7"

check_question("none",
               options = c("single quotation mark", "double quotation mark", "none", "space", "dot"), 
               type = "radio",
button_label = "Check answer",
random_answer_order = TRUE,
q_id = "Q6.7", 
right = "You are right: this DSV file does not have a delimiter character. This would be problematic if some of the values included commas but, as they don't, it's not a problem.",
wrong = "No, take a closer look at the extract of the DSV file printed above.")

```

:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

Import both data files from @DabrowskaExperienceAptitudeIndividual2019 using the `read.csv` function as described above. Save the first as the `R` object `L1.data` (as in the example above) and the second as `L2.data`. Then, answer the following questions.

[**Q6.8**]{style="color:green;"} In the two data files from @DabrowskaExperienceAptitudeIndividual2019, each row corresponds to one participant. How many L1 participants were included in this study?

```{r}
#| echo: false
#| label: "Q6.8"

check_question("90",
               options = c("90", "31", "221", "2790", "17"), 
               type = "radio",
button_label = "Check answer",
random_answer_order = TRUE,
q_id = "Q6.8", 
right = "That's right!",
wrong = "Hummm, are you sure?")
check_hint("Use the command `View(L1.data)` to display the full dataset in a new tab.",
           hint_title = "🐭 Click on the mouse for a hint.")

```

[**Q6.9**]{style="color:green;"} How many L2 participants were included in @DabrowskaExperienceAptitudeIndividual2019's study?

```{r}
#| echo: false
#| label: "Q6.9"

check_question("67",
               options = c("67", "45", "90", "220", "306"), 
               type = "radio",
button_label = "Check answer",
random_answer_order = TRUE,
q_id = "Q6.9", 
right = "That's right!",
wrong = "Hummm, are you sure?")
check_hint("How many rows does the dataframe `L2.data` contain?", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

[**Q6.10**]{style="color:green;"} Compare the data frames containing the L1 and L2 data. Which dataset contains more variables?

```{r}
#| echo: false
#| label: "Q6.10"

check_question("L2.data",
               options = c("L2.data", "L1.data", "They have the same number of variables"), 
               type = "radio",
button_label = "Check answer",
random_answer_order = FALSE,
q_id = "Q6.10", 
right = "That's right! The L2 dataset includes additional information such as the participants' native language(s) and the age they were when they first moved to an English-speaking country.",
wrong = "Hummm, are you sure?")
check_hint("Remember that, in these two data frames, each colum represents a variable. Use the `View()` function to display the data frames and compare them.", 
           hint_title = "🐭 Click on the mouse for a hint.")

```

[**Q6.11**]{style="color:green;"} You have saved the two datasets to your local `R` environment as `L1.data` and `L2.data`. What kind of `R` objects are `L1.data` and `L2.data`? You can find out by using the command `class()`. It simply takes the name of the object as its only argument.

```{r}
#| echo: false
#| label: "Q6.11"

check_question("data frame",
               options = c("data frame", "character", "integer", "tibble", "table", "list"), 
               type = "radio",
button_label = "Check answer",
q_id = "Q6.11", 
random_answer_order = TRUE,
right = "That's right!",
wrong = "That's incorrect.")
check_hint("Use the command `class(L1.data)` to find out which type of `R` object `L1.data` is.", hint_title = "🐭 Click on the mouse for a hint.")

```

[**Q6.12**]{style="color:green;"} Why does the L2 dataset contain the variable `NativeLg`, but not the L1 dataset?

```{r}
#| echo: false
#| label: "Q6.12"

check_question("Because, in this study, all L1 participants have English as their native language.",
               options = c("Because, in this study, all L1 participants have English as their native language.", "Because some or all of the L1 participants did not wish to answer this question.", "Because this variable was removed from the dataset for data protection reasons.", "Because Dr. Dąbrowska decided not to collect this information for L1 participants."), 
               type = "check",
button_label = "Check answer",
q_id = "Q6.12", 
random_answer_order = TRUE,
right = "That's right!",
wrong = "No, that's not the reason.")

```
:::
### Check your progress 🌟 {.unnumbered}

You have successfully completed [`r checkdown::insert_score()` out of 12 questions]{style="color:green;"} in this chapter.
