---
title: "Learning Quarto"
subtitle: "by reproducing the descriptive statistics of Dąbrowska's (2019) study"
author: "Elen Le Foll"
date: last-modified
format: pptx
bibliography: references.bib
csl: international-journal-of-learner-corpus-research.csl
---

# Introduction

The aim of this report is to reproduce the descriptive statistics reported in Dąbrowska (2019: 5-6) using the original datasets (Dąbrowska 2019: Appendices):

> **Method**
>
> **Participants**
>
> Ninety native speakers (42 male and 48 female) and 67 nonnative speakers of English (21 male and 46 female) were recruited through personal contacts, church and social clubs, and advertisements in local newspapers. Participants were told that the purpose of the study was to examine individual differences in native and nonnative speakers’ knowledge of English and whether these differences are related to their linguistic experience and abilities. All participants signed a written consent form before the research commenced.
>
> The L1 participants were all born and raised in the United Kingdom and were selected to ensure a range of ages, occupations, and educational backgrounds. The age range was from 17 to 65 years (M = 38, SD = 16). Twenty-two percent of the participants held manual jobs, 24% held clerical positions, and 28% had professional-level jobs or were studying for a degree; the remaining 26% were occupationally inactive (i.e., unemployed, retired, or homemakers). In terms of education, participants’ backgrounds ranged from no formal qualifications to Ph.D., with corresponding differences in the number of years spent in full-time education (from 10 to 21; M = 14, SD = 2). Six participants reported a working knowledge of another language; the rest described themselves as monolinguals.

```{r}
#| label: setup
#| include: false

library(here)
library(tidyverse)
```

```{r}
#| label: import-data
#| include: false
  
L1.data <- read.csv(file = here("data", "L1_data.csv"))
L2.data <- read.csv(file = here("data", "L2_data.csv"))
```

`r nrow(L1.data)` native speakers and `r nrow(L2.data)` nonnative speakers of English...

## A note on inline citations

To conduct this reproduction study, we used the data from @DabrowskaExperienceAptitudeIndividual2019. To find out more about using `R` for data science, @wickhamDataScienceImport2023 is recommended reading.

# References
