# Ch. 10: Tasks & Quizzes {.unnumbered}

::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q10.1**]{style="color:green;"} Which of the following labels can be added or modified using the `labs()` function?

```{r}
#| echo: false
#| label: "Q10.1"
library(checkdown)
check_question(c("x-axis label",
                 "y-axis label",
                 "plot title",
                 "plot subtitle", 
                 "plot caption",
                 "plot alt-text"),
                 options = c("x-axis label",
                 "y-axis label",
                 "plot title",
                 "plot subtitle", 
                 "plot caption",
                 "plot alt-text"), 
               type = "check", 
               random_answer_order = TRUE,
               q_id = "Q10.1", 
               alignment = "vertical",
               button_label = "Check answer",
               right = "That's right and there are many more possibilities! For instance, it's also possible to change the title of a plot's legend using the labs() function.",
               wrong = "No, there are more labels that can be ammended.")
check_hint("Check out the function's help file to find out more about its possibilities by entering ?labs in the Console.", hint_title = "🐭 Click on the mouse for a hint.")
```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q10.2**]{style="color:green;"} Study @fig-NativeLanguagesBarplot and think about which {ggplot2} functions were used to generate it. 🤔

Then, click on the "Show `R` code" button just above the figure to compare your intuitions with the actual code. Note that there may well be more than one solution, so do try out your version and see if you can spot any differences!
:::
::::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q10.3**]{style="color:green;"} Compare the two plots below. Which one makes it easier to see which occupational group has the fewest participants and why?

```{r}
#| include: false
library(here)
library(tidyverse)

Dabrowska.data <- readRDS(file = here("data", "processed", "combined_L1_L2_data.rds"))
```

:::: columns
::: {.column width="45%"}
**Bar plot**

```{r}
#| code-fold: true
#| code-summary: "Show `R` code for the bar plot"

Dabrowska.data |> 
  mutate(OccupGroup = fct_recode(OccupGroup,
                                 `Professionally inactive` = "I",
                                 `Clerical profession` = "C",
                                 `Manual profession` = "M",
                                 `Professional-level job/\nstudent` = "PS")) |> 
  filter(Gender == "M") |> 
  ggplot(mapping = aes(x = OccupGroup, 
                       fill = OccupGroup)) +
  geom_bar() +
  labs(x = NULL,
       y = "Participants") +
  scale_fill_viridis_d() +
  coord_flip() +
  theme_minimal() +
  theme(axis.text = element_text(size = 15),
        legend.position = "none")
```
:::

::: {.column width="8%"}
<!-- empty column to create gap -->
:::

::: {.column width="37%"}
**Pie chart**

```{r}
#| code-fold: true
#| code-summary: "Show `R` code for the pie chart"

Dabrowska.data |> 
  mutate(OccupGroup = fct_recode(OccupGroup,
                                 `Professionally inactive` = "I",
                                 `Clerical profession` = "C",
                                 `Manual profession` = "M",
                                 `Professional-level job/\nstudent` = "PS")) |> 
  filter(Gender == "M") |> 
  ggplot(mapping = aes(x = "", fill = OccupGroup)) +
    geom_bar(width = 1) +
    scale_fill_viridis_d(direction = -1) +
    coord_polar("y") +
    theme_void(base_size = 20) # This increases the font size.
```
:::
::::

```{r}
#| echo: false
#| label: "Q10.3"

check_question(c("The bar plot because humans are better at distinguishing small differences in lengths than in angles.",
                 "The bar plot because the labels are immediately next to the corresponding bars."),
               options = c("The bar plot because humans dislike round objects and prefer objects with sharp angles.",
                           "The bar plot because humans are better at distinguishing small differences in lengths than in angles.",
                           "The pie chart as most humans learn to read clocks early on in life.",
                           "The bar plot because the labels are immediately next to the corresponding bars.",
                           "The pie chart because the colours are easier to distinguish when they are close to each other."),
               type = "check",
               q_id = "Q10.3", 
               button_label = "Check answer",
               random_answer_order = TRUE,
               right = "Correct! Bar charts are better suited than pie charts because humans are much better at estimating lengths than angles (see Section 10.1.10), and the labels are immediately next to the corresponding bars here.",
               wrong = "No, try again.")
check_hint("Check out Section 10.1.10.", 
           hint_title = "🦉 Hover over the owl for a first hint.",
           type = "onmouseover")
check_hint("Two reasons are correct.", 
           hint_title = "🐭 Click on the mouse for a second hint.")

```
:::::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

Using the {ggplot2} library, create a bar plot that shows the distribution of occupational groups (`OccupGroup`) among male L1 and L2 participants in @DabrowskaExperienceAptitudeIndividual2019's study.

[**Q10.4**]{style="color:green;"} Drawing on the information provided by your bar plot, how many male participants reported having manual jobs?

```{r}
#| echo: false
#| label: "Q10.4"

check_question("exactly 20",
               options = c("under 10",
                           "exactly 12",
                           "exactly 20",
                           "exactly 30",
                           "a little more than 40"),
               type = "radio",
               q_id = "Q10.4", 
               button_label = "Check answer",
               right = "That's right! Your bar plot should make that immediately obvious.",
               wrong = "No, are you sure that you have plotted the distribution of occupational groups among male participants only?")

```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q10.4"
#| eval: false

Dabrowska.data |> 
  filter(Gender == "M") |> 
  ggplot(mapping = aes(x = OccupGroup)) +
  geom_bar() +
  labs(x = "Occupational group", 
       y = "Male participants") +
  theme_minimal()
```

[**Q10.5**]{style="color:green;"} Is the legend in the bar plot that you have created necessary?

```{r}
#| echo: false
#| label: "Q10.5"

check_question("No, it only adds clutter.",
               options = c("No, it only adds clutter.",
                           "Yes, it makes the plot easier to interpret.",
                           "Not really, but it makes the plot look more professional.",
                           "Not really, but it's impossible to remove the legend using the {ggplot2} library."),
               type = "radio",
               q_id = "Q10.5", 
               button_label = "Check answer",
               right = "That's right: the legend is redundant and does not make the plot easier to interpret. We can remove the legend by adding a theme layer that changes the position of the legend to \"none\". Click below to find out how to integrate this in ggplot2 code.",
               wrong = "Are you sure?")

```

```{r}
#| code-fold: true
#| code-summary: "See code."
#| eval: false
#| source-line-numbers: "7"

Dabrowska.data |> 
  filter(Gender == "M") |> 
  ggplot(mapping = aes(x = OccupGroup, 
                       fill = OccupGroup)) +
  geom_bar() +
  theme_minimal() +
  theme(legend.position = "none")
```

[**Q10.6**]{style="color:green;"} Create a pie chart that shows the distribution of occupational groups among male participants (as shown in @fig-OccupGroupPieChart). Which line of code is essential to create a pie chart using {ggplot2}?

```{r}
#| echo: false
#| label: "Q10.6"

check_question("coord_polar(\"y\")",
               options = c("coord_polar(\"y\")",
                           "theme_void()",
                           "coord(\"polar\")",
                           "geom_piechart()",
                           "geom_circle()"),
               type = "radio",
               q_id = "Q10.6", 
               button_label = "Check answer",
               right = "Yes, well done!",
               wrong = "Not quite. Try adding and removing these lines of code in your piechart script to see the effect. If you're stuck, check out the code answer below.")

```

```{r}
#| code-fold: true
#| code-summary: "Show code to create pie chart below."
#| fig-width: 4
#| fig-height: 2

Dabrowska.data |> 
  filter(Gender == "M") |> 
  ggplot(mapping = aes(x = "", 
                       fill = OccupGroup)) +
    geom_bar(width = 1) +
    coord_polar("y") +
    theme_void()
```

[**Q10.7**]{style="color:green;"} Transform the pie chart that you just created in [c.]{style="color:green;"} to make it look like @fig-PieChartRecode. To achieve this, wrangle the data *before* piping it into the `data` argument of the `ggplot()` function. Which {tidyverse} function can you use to rename the labels?

```{r}
#| echo: false
#| label: "Q10.7"

check_question("fct_recode()",
               options = c("fct_recode()",
                           "rename()",
                           "fct_mutate()",
                           "fct_relevel()",
                           "labs()"),
               type = "radio",
               q_id = "Q10.7", 
               random_answer_order = TRUE,
               button_label = "Check answer",
               right = "Absolutely!",
               wrong = "No, that won't do the trick. Re-read Section 10.1.6 for a quick recap.")

```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q10.7"
#| source-line-numbers: "1-6,13"
#| fig-width: 4
#| fig-height: 2
#| label: "fig-PieChartRecode"

Dabrowska.data |> 
  mutate(`OccupGroup` = fct_recode(OccupGroup,
                                 `Professionally inactive` = "I",
                                 `Clerical profession` = "C",
                                 `Manual profession` = "M",
                                 `Professional-level job` = "PS")) |> 
  filter(Gender == "M") |> 
  ggplot(mapping = aes(x = "", 
                       fill = OccupGroup)) +
  geom_bar(width = 1) +
  coord_polar("y") +
  theme_void() +
  labs(fill = "Occupational group") # This last line of code changes the title of the legend, which is the label for the variable associated with the `fill` aestetics. 
  
```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q10.13**]{style="color:green;"} Which line of code needs to be added to the code used to generate @fig-GrammarDensity to produce a two-panel density plot as in the figure below?

```{r}
#| echo: false
#| label: "Q10.13"

check_question("facet_wrap(~ Group)",
               options = c("facets(~ \"Group\")",
                           "fct_wrap(\"Group\")",
                           "facet_wrap(Group)",
                           "facet_wrap(~ Group)",
                           "geom_grid(Group, facet = 2)"),
               type = "radio",
               q_id = "Q10.13", 
               button_label = "Check answer",
               right = "You're right! Note that facet_wrap(\"Group\") also works and produces the same output.",
               wrong = "Not quite, try these lines out yourself and check out @sec-facets to find out more.")

```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to help you answer Q10.13."

ggplot(data = Dabrowska.data,
       mapping = aes(x = Grammar)) +
  geom_density(fill = "purple") +
  facet_wrap(~ Group) +
  labs(x = "Scores on English grammar test") +
  theme_bw()
```

[**Q10.14**]{style="color:green;"} Create four density plots to visualise the distribution of participants' `ART` (Author Recognition Test), `Blocks` (non-verbal IQ test), `Colloc` (English collocation test), and `Vocab` (English vocabularly test) scores.

Which variable's distribution is closest to a normal distribution?

```{r}
#| echo: false
#| label: "Q10.14"

check_question("Blocks",
               options = c("ART",
                           "Blocks",
                           "Colloc",
                           "Vocab",
                           "ART and Vocab"),
               type = "radio",
               q_id = "Q10.14", 
               button_label = "Check answer",
               right = "That's right. Even though the distribution of Blocks scores is not perfectly symmetrical, it is less skewed than that of the other three variables.",
               wrong = "No, check [Chapter 8](https://elenlefoll.github.io/RstatsTextbook/8_DescriptiveStats.html#sec-Normal) to find out more about normal and non-normal distributions.")

```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to help you answer Q10.13."
#| eval: false

Dabrowska.data |> 
  select(Vocab, Colloc, ART, Blocks) |>  
  tidyr::gather() |>  # This function from tidyr converts a selection of variables into two variables: a key and a value. The key contains the names of the original variable and the value the data. This means we can then use the facet_wrap function from ggplot2
  ggplot(aes(value)) +
    theme_bw() +
    facet_wrap(~ key, scales = "free", ncol = 2) +
    scale_x_continuous(expand=c(0,0)) +
    geom_density(fill = "purple")
```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q10.15**]{style="color:green;"} Create a boxplot to compare how participants in different occupational groups (`OccupGroup`) performed on the English grammar test. Which part of the code used to produce @fig-GrammarBoxPlotGroup do you need to modify to achieve this?

```{r}
#| echo: false
#| label: "Q10.15"

check_question("x = OccupGroup",
               options = c("x = OccupGroup",
                           "y = Grammar",
                           "Dabrowska.data",
                           "geom_boxplot()"),
               random_answer_order = TRUE,
               q_id = "Q10.15", 
               type = "radio",
               button_label = "Check answer",
               right = "Correct!",
               wrong = "No. Check out the code solution below if you're unsure.")
```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q10.15."
#| eval: false
#| fig-width: 8
#| fig-height: 4
#| source-line-numbers: "3"

Dabrowska.data |>
  ggplot(mapping = aes(y = Grammar, 
                       x = OccupGroup)) +
  geom_boxplot() +
  theme_minimal() +
  labs(y = "Grammar scores")
```

[**Q10.16**]{style="color:green;"} The code below was used to create @fig-OccupBoxPlot, except that the arguments of the `aes()` function have been deleted. Which data mappings were specified inside the `aes()` function to produce @fig-OccupBoxPlot?

```{r}
#| eval: false

Dabrowska.data |>
  mutate(Group = fct_recode(Group,
                            `L1 participants` = "L1",
                            `L2 participants` = "L2")) |> 
  ggplot(mapping = aes(█ █ █ █ █ █ █ █ █ █ █ █)) +
    geom_boxplot(alpha = 0.8) +
    scale_fill_viridis_d(option = "viridis") +
    scale_y_continuous(breaks = seq(0, 100, 10)) +
    labs(y = "Grammar scores", 
         x = NULL,
         fill = "Occupational group",
         title = "English grammar comprehension test results",
         subtitle = "among L1 and L2 participants with different occupations") +
    theme_bw() +
    theme(element_text(size = 12),
          legend.position = "bottom", # We move the legend to the bottom of the plot. 
          legend.box.background = element_rect()) # We add a frame around the legend.
```

```{r}
#| echo: false
#| label: "fig-OccupBoxPlot"
#| fig-alt: "The previous box plot has been subdivided further according to participant occupation, producing 4 box plots each. The box plots for the L1 speakers are quite similar to each other and have smaller and greater ranges, whereas the L2 plots have larger ranges, with the exception of the 'I' occupation group, which is small and falls entirely between 90 and 100."

Dabrowska.data |>
  mutate(Group = fct_recode(Group,
                            `L1 participants` = "L1",
                            `L2 participants` = "L2")) |> 
  ggplot(mapping = aes(y = Grammar, 
                       x = Group, 
                       fill = OccupGroup,
                       facet = OccupGroup)) +
    geom_boxplot(alpha = 0.8) +
    scale_fill_viridis_d(option = "viridis") +
    scale_y_continuous(breaks = seq(0, 100, 10)) +
    labs(y = "Grammar scores", 
         x = NULL,
         fill = "Occupational group",
         title = "English grammar comprehension test results",
         subtitle = "among L1 and L2 participants with different occupations") +
    theme_bw() +
    theme(element_text(size = 12),
          legend.position = "bottom", # We move the legend to the bottom of the plot. 
          legend.box.background = element_rect()) # We add a frame around the legend.

```

```{r}
#| echo: false
#| label: "Q10.16"

check_question(c("y = Grammar", 
                 "x = Group",
                 "fill = OccupGroup",
                 "facet = OccupGroup"),
               options = c("y = Grammar",
                           "y = OccupGroup",
                           "x = OccupGroup",
                           "x = Group",
                           "legend = OccupGroup",
                           "colour = OccupGroup",
                           "fill = OccupGroup",
                           "facet = Group",
                           "facet = OccupGroup"),
               type = "check",
               q_id = "Q10.16", 
               random_answer_order = TRUE,
               button_label = "Check answer",
               right = "That's right! 🎉 Well done, that wasn't easy!",
               wrong = "No, not quite. Try out the code with these different mappings to see what effect they have.")
check_hint("Copy the code and experiment with various combinations of the proposed mappings.", 
           hint_title = "🦉 Hover over the owl for a first hint.",
           type = "onmouseover")
check_hint("This boxplot requires a total of four data mappings.", 
           hint_title = "🐭 Click on the mouse for a second hint.")
```

 
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

In this quiz, you will explore another hypothesis:

> L2 speakers' grammar comprehension scores (`Grammar`) are positively correlated with their length of residence in the UK (`LoR`): L2 speakers who have lived in the UK for longer have a better understanding than those who arrived more recently.

Using {ggplot2}, create a **scatter plot** that allows you to explore this hypothesis.

[**Q10.17**]{style="color:green;"} What do you need to do before piping the data into the `ggplot()` function?

```{r}
#| echo: false
#| label: "Q10.17"

check_question("filter(Group == \"L2\")",
               options = c("filter(Group == \"L2\")",
                           "filter(Group != \"L2\")",
                           "select(Group, \"L2\")",
                           "select(LoR, L2)",
                           "select(Grammar, LoR, Group, L2)"),
               random_answer_order = TRUE,
               q_id = "Q10.17", 
               type = "radio",
               button_label = "Check answer",
               right = "✅",
               wrong = "❌ No, that's not it.")
check_hint("You'll find the answer in [Section 10.1.4](https://elenlefoll.github.io/RstatsTextbook/10_Dataviz.html#sec-ggplotData).", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q10.17."
#| eval: false

Dabrowska.data |> 
  filter(Group == "L2") |> 
  ggplot(mapping = aes(x = LoR, 
                       y = Grammar)) +
  geom_point() +
  labs(x = "Length of residence in the UK (years)",
       y = "Grammar comprehension test scores") +
  theme_bw()
```

[**Q10.18**]{style="color:green;"} On your scatter plot, which data points are clearly outliers?

```{r}
#| echo: false
#| label: "Q10.18"

check_question("The L2 speaker who has lived in the UK for more than 40 years.",
               options = c("The L2 speaker who has lived in the UK for more than 40 years.",
                           "The L2 speaker who scored 0 on the grammar test.",
                           "The three L2 speakers who scored 100 on the grammar test.",
                           "All the L2 speakers who have lived in the UK for 15 years or more.",
                           "The 90 speakers with missing values."),
               random_answer_order = TRUE,
               q_id = "Q10.18", 
               type = "check",
               button_label = "Check answer",
               right = "✅",
               wrong = "❌ No, that's not it.")
check_hint("If you are getting a warning message telling you that ggplot2 has \"removed 90 rows containing missing values or values outside the scale range\", you have not filtered the data to exclude L1 speakers before plotting the data. L1 speakers all have \"NA\" values for the variable LoR (corresponding to the length of residence in the UK).", 
           hint_title = "🐭 Click on the mouse for a hint.")

```

Identifying outliers is important to check and understand our data. If we spot an outlier corresponding to a participant having lived in a country for 400 years, we can immediately tell that something has gone wrong during data collection and/or pre-processing. In this case, however, common sense tells us that it is prefectly possible to have lived 40+ years in a country. Still, compared to the rest of the data, the outlier is striking and we should check that it is not due to an error.

[**Q10.19**]{style="color:green;"} Using data wrangling functions from the {tidyverse} (see @sec-DataWrangling), check whether the outlier participant is old enough to have lived in the UK for longer than 40 years. How old were they when they first came to the UK?

```{r}
#| echo: false
#| label: "Q10.19"

check_question(c("20", "twenty"),
               button_label = "Check answer",
               q_id = "Q10.19", 
               right = "✅ Good job!" ,
               wrong = "❌ That's not it. Check the hint below, or the sample code if you're stuck.")
check_hint("To answer this question, you need to filter the data to only retain L2 participants who have lived in the UK for longer than 40 years. Then you can use the select() function to output only their age and length of residence.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q10.19."
#| eval: false

Dabrowska.data |> 
  filter(LoR > 40) |> 
  select(Age, LoR)

62-42
```

[**Q10.20**]{style="color:green;"} In order to better visualise a potential correlation between L2 participants' grammar comprehension test results and how long they've lived in the UK, we will exclude the outlier and re-draw the scatter plot. Which function can you use to exclude the identified outlier before piping the data into the `ggplot()` function?

```{r}
#| echo: false
#| label: "Q10.20"

check_question("filter(LoR < 40)",
               options = c("filter(LoR < 40)",
                           "filter(LoR >= 40)",
                           "filter(LoR != 40)",
                           "filter(LoR == \"< 40\")",
                           "filter(LoR ~ 40)"),
               random_answer_order = TRUE,
               q_id = "Q10.20", 
               type = "radio",
               button_label = "Check answer",
               right = "✅ That's right, well done!",
               wrong = "❌ No, that's not the one. Try them out and see how they change your scatterplot.")
check_hint("To answer this question, it's worth revising relational operators (see [Section 5.5](https://elenlefoll.github.io/RstatsTextbook/5_GettingStaRted.html#sec-RelationalOperators)).", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q10.20."
#| eval: false

Dabrowska.data |> 
  filter(Group == "L2") |> 
  filter(LoR < 40) |> 
  ggplot(mapping = aes(x = LoR, 
                       y = Grammar)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  labs(x = "Length of residence in the UK (years)",
       y = "Grammar comprehension test scores") +
  theme_bw()
```

[**Q10.21**]{style="color:green;"} Having removed the outlier point from the scatter plot, now add a regression line to visualise the correlation between the two variables. Which of the following options best fits the gap in the sentence below?

> In this group of L2 speakers, there is \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_ correlation between participants' grammar test scores and their length of residence in the UK.

```{r}
#| echo: false
#| label: "Q10.21"

check_question("a positive",
               options = c("a positive",
                           "a negative",
                           "no",
                           "a perfect"),
               random_answer_order = TRUE,
               q_id = "Q10.21", 
               button_label = "Check answer",
               type = "radio",
               right = "✅",
               wrong = "❌")
check_hint("To answer this question, you should ask yourself: Does the regression line draw an upward or a downward slope? How steep is the slope? And how close are the points to the line?).", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q10.21."
#| eval: false

Dabrowska.data |> 
  filter(Group == "L2") |> 
  filter(LoR < 40) |> 
  ggplot(mapping = aes(x = LoR, 
                       y = Grammar)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  labs(x = "Length of residence in the UK",
       y = "Grammar comprehension test scores") +
  theme_bw()
```
:::
### Check your progress 🌟 {.unnumbered}

Well done! You have successfully completed this chapter on data visualisation using the Grammar of Graphics approach. You have answered [`r checkdown::insert_score()` out of 21 questions]{style="color:green;"} correctly.
