# Ch. 11: Tasks & Quizzes {.unnumbered}

::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q11.1**]{style="color:green;"} Which of the following statements best describes random sampling?

```{r}
#| echo: false
#| label: "Q11.1"

library(checkdown)

check_question("Each individual in the population has the same known probability of being chosen (e.g. 0.001%).",
  options = c(
    "Every member of the target population is guaranteed to take part in the study.",
    "Participants are selected because they are easy to reach (e.g. university students).",
    "Each individual in the population has the same known probability of being chosen (e.g. 0.001%).",
    "The sample is built so that its composition (age, gender, socioeconomic status, etc.) matches known census figures."
  ),
  random_answer_order = TRUE,
  q_id = "Q11.1", 
  type                = "radio",      
  button_label        = "Check answer",
  right = "Correct! Random sampling (also called probability sampling) gives every individual an equal, known chance of being selected.",
  wrong = "Incorrect. Review the definition of random sampling in the table above. It is not about convenience or quota‑matching.")
```

[**Q11.2**]{style="color:green;"} True or false: In a convenience sample, the researcher can ensure that the sample is representative of the entire population if they are careful to match the sample to census demographics (age, gender, region, etc.).

```{r}
#| echo: false
#| label: "Q11.2"

check_question("False",
  options = c(
    "True",
    "False"),
  type                = "radio",
  q_id = "Q11.2", 
  button_label        = "Check answer",
  right = "That's right. Even when a convenience sample is \"quota‑matched\" on observable demographics, it remains a convenience sample because participants first volunteered to take part (e.g. they signed up on a platform like MTurk). This means that unobserved factors, such as motivation, socioeconomic status, and language background, can still render the sample unrepresentative, thus limiting the external validity of inferential data analyses made on the basis of these data.",
  wrong = "Incorrect.")
```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

Consider the following research question:

-   Across all adult English speakers in the UK, do L1 speakers, on average, achieve higher scores than L2 speakers on the English receptive vocabulary test?

[**Q11.3**]{style="color:green;"} Which null hypothesis could you formulate for this research question?

```{r}
#| echo: false
#| label: "Q11.3"

check_question("Across the entire adult UK population, there is no difference in the mean scores of L1 vs. L2 speakers on the English receptive vocabulary test.",
               options = c("Across the entire adult UK population, there is no difference in the mean scores of L1 vs. L2 speakers on the English receptive vocabulary test.",
                           "Across the entire adult UK population, the mean scores of both L1 and L2 speakers on the English receptive vocabulary test are null.",
                           "Across the entire adult UK population, there is no statistically significant difference between the mean scores of L1 speakers and those of L2 speakers on the English receptive vocabulary test."),
               random_answer_order = TRUE,
               q_id = "Q11.3",
               type = "radio",
               button_label = "Check answer",
               right = "That's right, well done! 🎉",
               wrong = "No, that's incorrect.")
```

[**Q11.4**]{style="color:green;"} Run a *t*-test on the @DabrowskaExperienceAptitudeIndividual2019 data to test the null hypothesis that you selected in Q11.3 above. What is the value of the *t*-statistic?

```{r}
#| echo: false
#| label: "Q11.4"

check_question("4.6768",
               options = c("4.6768",
                           "133.83",
                           "0.000007032",
                           "9.425801",
                           "23.240497"),
               random_answer_order = TRUE,
               q_id = "Q11.4", 
               type = "radio",
               button_label = "Check answer",
               right = "Correct!",
               wrong = "Incorrect.")
check_hint("As a reminder, the results of the English receptive vocabulary test are stored in the `Vocab` variable and the distinction between the L1 and the L2 participants' results is made in the `Group` variable so these are the two variables that will need to be entered in the `t.test()` formula.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| include: false
library(here)
library(tidyverse)
library(effectsize)

Dabrowska.data <- readRDS(file = here("data", "processed", "combined_L1_L2_data.rds"))

L1.data <- Dabrowska.data |> 
  filter(Group == "L1")

L2.data <- Dabrowska.data |> 
  filter(Group == "L2")
```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q11.4."
#| eval: false

# It is always a good idea to visualise the data before running a statistical test:
Dabrowska.data |> 
  ggplot(mapping = aes(y = Vocab,
                       x = Group)) +
  geom_boxplot(alpha = 0.7) +
  theme_bw()

# Running the t-test:
t.test(formula = Vocab ~ Group, 
       data = Dabrowska.data)

```

[**Q11.5**]{style="color:green;"} What is the *p*-value associated with the *t*-test you ran above?

```{r}
#| echo: false
#| label: "Q11.5"

check_question("0.000007032",
               options = c("7.03206",
                           "7.032 multiplied by minus six",
                           "7.032 to the power of minus six",
                           "0.0000007032",
                           "0.000007032"),
               random_answer_order = TRUE,
               q_id = "Q11.5", 
               type = "radio",
               button_label = "Check answer",
               right = "Correct!",
               wrong = "Incorrect.")
```

[**Q11.6**]{style="color:green;"} The *p*-value is very, very small. What does this mean? Select all that apply.

```{r}
#| echo: false
#| label: "Q11.6"

check_question("If there were no association between English native-speaker status and receptive English vocabulary scores, it is very unlikely that we would observe this difference in mean values by chance alone.",
               options = c("If there were no association between English native-speaker status and receptive English vocabulary scores, it is very unlikely that we would observe this difference in mean values by chance alone.",
                           "The true difference in the mean receptive English vocabulary scores of English native and non-native speakers is not equal to zero.",
                           "The true difference (across the full population) in the mean receptive English vocabulary scores of English native speakers is only very slightly higher than that of non-native speakers.",
                           "Given the sample size, the results of this t-test are highly precise.",
                           "The difference in the mean receptive English vocabulary scores of English native and non-native speakers is highly relevant."),
               random_answer_order = TRUE,
               q_id = "Q11.6", 
               type = "check",
               button_label = "Check answer",
               right = "Correct! 🎉",
               wrong = "Incorrect.")
check_hint("Only one of these statements is correct.", 
           hint_title = "🐭 Click on the mouse for a hint.")

```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

Compute three Cohen's *d* values to capture the magnitude of the difference between L1 and L2 speakers:

1.  English grammar (`Grammar`) test scores
2.  English receptive vocabulary (`Vocab`) test scores, and
3.  English collocation (`Colloc`) test scores.

As these are three standardised effect sizes, we can compare them.

[**Q11.7**]{style="color:green;"} Comparing the three Cohen's *d* that you computed, which statement(s) is/are true?

```{r}
#| echo: false
#| label: "Q11.7"

check_question("By far the largest difference between L1 and L2 speakers is in the English collocation test.",
               options = c("By far the largest difference between L1 and L2 speakers is in the English collocation test.",
                           "The largest difference between L1 and L2 speakers is in the receptive English vocabulary test.",
                           "The effect sizes across the three English tests are practically equal.",
                           "The difference in all three effect sizes is statistically significant.",
                           "The smallest effect size is observed in the collocation test."),
               random_answer_order = TRUE,
               q_id = "Q11.7", 
               type = "check",
               button_label = "Check answer",
               right = "That's right, well done! 🎉",
               wrong = "No, that's incorrect.")
check_hint("Only one of these statements is correct.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q11.7."
#| eval: false

cohens_d(Grammar ~ Group, 
       data = Dabrowska.data)

cohens_d(Vocab ~ Group, 
       data = Dabrowska.data)

cohens_d(Colloc ~ Group, 
       data = Dabrowska.data)
```

[**Q11.8**]{style="color:green;"} Now compute 99% confidence intervals (CI) around the same three Cohen's *d* values. Which statement(s) is/are true?

```{r}
#| echo: false
#| label: "Q11.8"

check_question(c("All three differences in means are statistically significant at the $\alpha$-level of 0.01.",
                 "For all three effect sizes, the 99% CIs are wider than the 95% CI."),
               options = c("All three differences in means are statistically significant at the $\alpha$-level of 0.01.",
                 "All three differences in means are statistically confident at the $\alpha$-level of 0.01.",
                           "We can be 99% confident that all three differences in means are real.",
                           "In 99% of cases, L1 speakers perform better than L2 speakers on all three tests.",
                           "The most relevant difference can be found in the collocation test data.",
                           "For all three effect sizes, the 99% CIs are wider than the 95% CI."),
               random_answer_order = TRUE,
               q_id = "Q11.8", 
               type = "check",
               button_label = "Check answer",
               right = "Correct!",
               wrong = "No, not quite. Two of these statements are correct and you will need to select both to get this question right.")
check_hint("Two of these statements are correct. Click on the link below for a reminder as to how to compute 99% confidence intervals around Cohen's d.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show code to answer Q11.8."
#| eval: false

cohens_d(Grammar ~ Group, 
       data = Dabrowska.data,
       ci = 0.99)

cohens_d(Vocab ~ Group, 
       data = Dabrowska.data,
       ci = 0.99)

cohens_d(Colloc ~ Group, 
       data = Dabrowska.data,
       ci = 0.99)
```

[**Q11.9**]{style="color:green;"} Consider the code and its outputs below. We are now comparing the English grammar comprehension test scores of male and female native speakers of Romance languages only. Hence, we are now looking at a very small sample size comprising just five female and one male participants.

```{r}
Dabrowska.Romance <- Dabrowska.data |> 
  filter(NativeLgFamily == "Romance")

table(Dabrowska.Romance$Gender)

cohens_d(Grammar ~ Gender,
         data = Dabrowska.Romance)
```

Which statement(s) is/are true about the standardised effect size computed for the difference in the English comprehension grammar scores of male and female Romance L1 speakers in this very small dataset (n = 6)?

```{r}
#| echo: false
#| label: "Q11.9"

check_question(c("The effect size is negative, which means that the single male participant performed better on the grammar test than the five female participants.",
                           "According to Cohen's rule of thumb, the effect size is very large.",
                 "The difference in mean grammar test scores across the two genders is not statistically significant at the $\alpha$-level of 0.05.",
               "We cannot reject the null hypothesis that there is no difference between male and female performance on the grammar test."),
               options = c("All six participants did equally well at the $\alpha$-level of 0.05.",
                           "The effect size is negative, which means that the single male participant performed better on the grammar test than the five female participants.",
                           "According to Cohen's rule of thumb, the effect size is very large.",
                           "According to Cohen's rule of thumb, the effect size is very small.",
                           "The 95% CI around Cohen's d includes zero, which means that there is no practical difference between the performance of male and female participants on the grammar test.",
                           "We cannot reject the null hypothesis that there is no difference between male and female performance on the grammar test.",
                           "The difference in mean grammar test scores across the two genders is not statistically significant at the $\alpha$-level of 0.05."),
               random_answer_order = TRUE,
               q_id = "Q11.9", 
               type = "check",
               button_label = "Check answer",
               right = "That's right, very well done! 🎉 The results of this test demonstrate the value of statistical significance testing: Here, the difference between the male and female participants is very large but, because we have such a small sample size, the confidence interval of the effect size includes zero. This informs us that this large difference could very well be due to chance and not actually correspond to a real difference in the population.",
               wrong = "No, not quite. To get this question right, you need to select all four correct statements.")
check_hint("Here we have an interesting case of a large effect size, but with a very wide 95% confidence interval that includes zero. What does this mean?", 
           hint_title = "🦉 Hover over the owl for a first hint.",
           type = "onmouseover")
check_hint("Four of these statements are correct. And remember that no statistical significance test can tell us anything about whether an effect is \"meaningful\" or \"practical\". To make such a statement would require a lot of additional contextual information that we do not have access to.", 
           hint_title = "🐭 Click on the mouse for a second hint.")
```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

Generate a facetted scatter plot (similar to @fig-YearEducationPLotGrayBand) visualising the correlation between age (`Age`) and English receptive vocabulary test scores (`Vocab`) for L1 and L2 participants in two separate panels. Add a grey band visualising 95% confidence intervals around the regression lines.

[**Q11.10**]{style="color:green;"} Looking at your plot only, which of these statements can you confidently say is/are true?

```{r}
#| echo: false
#| label: "Q11.10"

check_question(c("There is more variability around the correlation estimate of the L2 than the L1 group.",
                 "In the L1 group, the correlation is positive, whereas in the L2 group it is negative."),
               options = c("Even though one correlation is positive and the other negative, in absolute values, the correlations are equally strong.",
                           "There is more variability around the correlation estimate of the L2 than the L1 group.",
                           "There is more variability around the correlation estimate of the L1 than the L2 group.",
                           "In the L1 group, the correlation is positive, whereas in the L2 group it is negative.",
                           "For both the L1 and L2 groups, the correlations are positive."),
               random_answer_order = TRUE,
               q_id = "Q11.10", 
               type = "check",
               button_label = "Check answer",
               right = "That's right, well done!",
               wrong = "No, that's not quite right. Two of these statements are correct. Can you find both?")
check_hint("Consider the direction of the regression lines and the width of the confidence bands around the lines.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show code to generate the plot needed to answer Q11.10."
#| eval: false

Dabrowska.data |> 
   ggplot(mapping = aes(x = Age, 
                        y = Vocab)) +
     geom_point() +
     facet_wrap(~ Group) +
     geom_smooth(method = "lm", 
                 se = TRUE) +
  theme_bw()
```

Now use the `cor.test()` function to find out how strong the correlations are in both the L1 and the L2 groups. Use the default $\alpha$-level of 0.05.

[**Q11.11**]{style="color:green;"} Based on the outputs of the `cor.test()` function, which of these statements can you confidently say is/are true? Note that, in this question and the next one, values are reported to two decimal places.

```{r}
#| echo: false
#| label: "Q11.11"

check_question(c("The correlation of Age and Vocab scores in the L1 sample is 0.37.",
                           "Based on the data available to us and an $\alpha$-level of 0.05, we cannot reject the null hypothesis of no correlation between Age and Vocab scores for the L2 population.",
                           "Assuming that the null hypothesis of no correlation between Age and Vocab scores in the L1 population is true, there is only a 0.03% chance of observing a correlation of 0.37 or larger in a sample of this size.",
                           "In absolute values, the correlation between Age and Vocab scores is stronger in the L1 than in the L2 sample."),
               options = c("The correlation of Age and Vocab scores in the L1 sample is 0.37.",
                 "The correlation of Age and Vocab scores in the L2 population is -0.20.",
                           "Based on the data available to us and an $\alpha$-level of 0.05, we cannot reject the null hypothesis of no correlation between Age and Vocab scores for the L2 population.",
                           "Assuming that the null hypothesis of no correlation between Age and Vocab scores in the L1 population is true, there is only a 0.03% chance of observing a correlation of 0.37 or larger in a sample of this size.",
                           "In absolute values, the correlation between Age and Vocab scores is stronger in the L1 than in the L2 sample."),
               random_answer_order = TRUE,
               q_id = "Q11.11", 
               type = "check",
               button_label = "Check answer",
               right = "That's right, well done! 🎉",
               wrong = "No, not quite. Do take another minute to carefully re-read all the statements. Four of them are correct.")
check_hint("Absolute values correspond to numeric values ignoring their sign, e.g. 3 and -3 correspond to the same absolute value of 3 and in absolute values, -10 is greater than 8.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show code to answer Q11.11."
#| eval: false

cor.test(formula = ~ Age + Vocab,
         data = L1.data)

cor.test(formula = ~ Age + Vocab,
         data = L2.data)
```

Consider the code and its output below. @fig-VocabArrivalPLot visualises the correlation between L2 speakers' English receptive vocabulary test scores (`Vocab`) and their age of arrival in the UK (`Arrival`).

```{r}
#| message: false
#| label: "fig-VocabArrivalPLot"
#| fig-cap: "Correlation between L2 speakers’ English vocabulary test scores and their age of arrival"
#| fig-alt: "A scatter plot showing the relationship between the L2 speakers' vocabulary test scores and their age of arrival in the UK. Each black dot represents an individual’s score. A blue regression line in the plot shows a negative trend: higher test scores are associated with younger ages of arrival. A shaded gray band sloping downwards is drawn around the blue line."

L2.data |> 
   ggplot(mapping = aes(x = Arrival, 
                        y = Vocab)) +
     geom_point() +
     geom_smooth(method = "lm", 
                 se = TRUE) +
  theme_bw()
```

[**Q11.12**]{style="color:green;"} Run a correlation test on the correlation visualised in @fig-VocabArrivalPLot. How strong is the correlation (to two decimal places)?

```{r}
#| echo: false
#| label: "Q11.12"

check_question(c("-0.31", "- 0.31", "-0.30", "- 0.30"),
               q_id = "Q11.12", 
               button_label = "Check answer",
               right = "That's right, very well done! ✅",
               wrong = "No. Are you perhaps looking at the wrong dataset (we are interested in L2 speakers only here) or the wrong variables (we are now looking at the association between Vocab and Arrival)?")
check_hint("Remember to round the correlation to two decimal places and to use a decimal point, not a comma.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show code to answer Q11.12."
#| eval: false

cor.test(formula = ~ Vocab + Arrival,
         data = L2.data)
```

[**Q11.13**]{style="color:green;"} How likely are we to observe such a strong correlation or an even stronger one in a sample of this size, if there is actually no correlation in the full L2 population?

```{r}
#| echo: false
#| label: "Q11.13"

check_question("about 1%",
               options = c("about 1%", 
                           "about 0.01%", 
                           "about 10%", 
                           "It's impossible to tell because the correlation is not statistically significant."),
               button_label = "Check answer",
               q_id = "Q11.13", 
               type = "radio",
               right = "That's right, well understood! 🎉",
               wrong = "No. You will need to interpret the p-value to answer this question.")
check_hint("A p-value of 1 would correspond to 100% likelihood, whereas a p-value of 0 would correspond to 0% probability of observing a correlation of this strength or larger in a sample of this size under the assumption of no correlation in the population.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```
:::
::: {.callout-caution collapse="true"}
#### [Optional (fun) task]{.unnumbered style="color:green;"} 🦖

Inspired by the Anscombe Quartet, @matejkaSameStatsDifferent2017 created a set of 12 pairs of variables that have the same descriptive statistics as the data that produce a scatter plot representing a tyrannosaurus (see @fig-DinoPLot).

```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 6
#| label: "fig-DinoPLot"
#| fig-cap: "Scatter plot depicting a tyrannosaurus (originally created by [Alberto Cairo](https://web.archive.org/web/20240620205540/http://www.thefunctionalart.com/2016/08/download-datasaurus-never-trust-summary.html))"
#| fig-alt: "A scatter plot with the x-axis ranging from 20 to 100, and the y-axis ranging from 0 to 100. The data points form a pattern that looks like a tyrannosaurus."

library(datasauRus)

datasaurus_dozen |> 
  filter(dataset == "dino") |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  theme_bw()
```

[**1.**]{style="color:green;"} Install and load the `datasauRus` package to access the `R` data object `datasaurus_dozen`:

```{r}
#| eval: false
install.packages("datasauRus")
library(datasauRus)

datasaurus_dozen
```

[**2.**]{style="color:green;"} Run the following code to compare the descriptive statistics of all three datasets within the `datasaurus_dozen`. The fourth set, `dino`, is the one visualised in @fig-DinoPLot. Note that there is a very small, negative correlation between all pairs of variables of `-0.0656`.

```{r}
datasaurus_dozen |> 
  group_by(dataset) |> 
  summarise(
    mean_x = mean(x), 
    mean_y = mean(y), 
    std_dev_x = sd(x), 
    std_dev_y = sd(y), 
    corr_x_y = cor(x, y))
```

[**3.**]{style="color:green;"} Run the following code to visualise the relationships between the other 12 pairs of variables.

```{r}
#| eval: false

ggplot(datasaurus_dozen, 
       aes(x = x, y = y, colour = dataset)) + 
  geom_point() + 
  facet_wrap(~ dataset, ncol = 3) +
  theme_bw() + 
  theme(legend.position = "none") + 
  labs(x = NULL,
       y = NULL)
```

[**4.**]{style="color:green;"} Add a `geom_` layer (see @sec-geoms) to the following {ggplot2} code to add a blue linear regression line in all 13 panels.

```{r}
#| code-fold: true
#| code-summary: "View code to complete Part 4 of the task."
#| eval: false
#| source-line-numbers: "4:6"

ggplot(datasaurus_dozen, 
       aes(x = x, y = y, colour = dataset)) + 
  geom_point() + 
  facet_wrap(~ dataset, ncol = 3) +
  geom_smooth(method = "lm",
              se = FALSE,
              colour = "blue") +
  theme_bw() + 
  theme(legend.position = "none") + 
  labs(x = NULL,
       y = NULL)  
```

Clearly, if we had only used correlation statistics to describe the relationships between these 13 pairs of variables, we would have missed some literally dinosaur-sized patterns! 🤯
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

The dataset from @DabrowskaExperienceAptitudeIndividual2019 includes the results of numerous additional tests that we have not yet examined.

In this task, we imagine that a friend of yours has decided to test whether, on average, male and female speakers performed equally well on four additional English grammar tests. The four variables that he selected correspond to participants' scores on English tests assessing participants' mastery of:

-   the active voice (`Active`)
-   the passive voice (`Passive`)
-   post-modified subjects (`Postmod`)
-   subject relatives (`SubRel`)

Your friend conducted the following four *t*-tests to find out whether he could reject the the null hypothesis of no difference in the average performance of male and female English speakers in these four grammar tests. He chose 0.05 as his significance threshold and, on the basis of these four tests, concluded that he could reject the null hypothesis in two out of four tests: the one concerning the active voice and the one about subject relatives.

```{r}
#| eval: false

t.test(formula = Active ~ Gender, 
       data = Dabrowska.data)

t.test(formula = Passive ~ Gender, 
       data = Dabrowska.data)

t.test(formula = Postmod ~ Gender, 
       data = Dabrowska.data)

t.test(formula = SubRel ~ Gender, 
       data = Dabrowska.data)
```

[**Q11.14**]{style="color:green;"} Your friend tells you that, based on the results of the two significant *t*-tests, he has decided to write a term paper focussing on how men have a better understanding of the active voice and subject relative constructions than women. You warn him that this approach reminds you of a questionable research practice (QRP) that you read about in the [FORRT glossary](https://forrt.org/glossary/english/). Which one?

```{r}
#| echo: false
#| label: "Q11.14"

check_question("HARKing",
                 options = c("HARKing",
                 "Salami slicing",
                 "Optional stopping",
                 "Creative use of outliers"), 
               type = "radio", 
               q_id = "Q11.14", 
               random_answer_order = TRUE,
               button_label = "Check answer",
               right = "That's right! This sounds like a typical case of HARKing (Hypothesizing After the Results are Known), which means presenting a hypothesis that is based on or informed by your results in a research report as if it was, in fact, a hypothesis that you came up with before looking at the data.",
               wrong = "No, that's not the problem, here.")
check_hint("You will need to go to the FORRT glossary to find out what these terms mean.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

[**Q11.15**]{style="color:green;"} You explain to your friend that, if he wants to conduct four separate tests on the same dataset, he needs to correct the *p*-values for multiple testing. Which reason(s) can you use to explain this to your friend?

```{r}
#| echo: false
#| label: "Q11.15"

check_question(c("Running multiple tests on the same dataset increases your chances of obtaining a false positive result.",
                 "Even though you chose an α-level of 0.05 for each test, your overall risk of erroneously rejecting the null hypothesis in at least one of these four tests is actually much higher than 5%."),
                 options = c("Running multiple tests on the same dataset increases your chances of obtaining a false positive result.",
                 "Even though you chose an α-level of 0.05 for each test, your overall risk of erroneously rejecting the null hypothesis in at least one of these four tests is actually much higher than 5%.",
                 "Running multiple tests on the same dataset makes each individual test weaker and less reliable.",
                 "The sample size gets split up when you conduct multiple tests, which results in less accurate results.",
                 "Multiple tests on the same dataset make it harder to detect real differences when they actually exist."), 
               type = "check", 
               q_id = "Q11.15", 
               random_answer_order = TRUE,
               alignment = "vertical",
               button_label = "Check answer",
               right = "Well done, your friend will be well-informed! 🤗",
               wrong = "No, not quite. You will need to select both correct options to get this question right.")
check_hint("Two of these reasons are correct. The other options correspond to common misconceptions about statistical signifiance testing.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

[**Q11.16**]{style="color:green;"} How high is your friend's risk of erroneously rejecting the null hypothesis in at least one of his four tests?

```{r}
#| echo: false
#| label: "Q11.16"

check_question("19%",
                 options = c("5%", "7%", "19%", "20%", "almost 100%"), 
               type = "radio", 
               q_id = "Q11.16", 
               alignment = "vertical",
               button_label = "Check answer",
               right = "That's right and that's definitely more than the 5% he originally intended to take!!",
               wrong = "❌")
check_hint("Section 11.7 showed how to calculate this risk for 20 individual tests. Can you adapt the formula to four tests?", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q11.16"
#| eval: false

1 - (1 - 0.05)^4 
```

[**Q11.17**]{style="color:green;"} Use Holm's correction to correct the four *p*-values that your friend obtained to account for the fact that he conducted four tests on the same dataset. After correction, which null hypotheses can be rejected at $\alpha$-level = 0.05?

```{r}
#| echo: false
#| label: "Q11.17"

check_question("Only the one concerning the active voice.",
                 options = c("All four of them.",
                             "Only the one concerning the active voice.",
                             "Only the one concerning the passive voice.",
                             "The one concerning the active voice and the one concerning subject relatives.",
                             "None of them."), 
               type = "radio", 
               q_id = "Q11.17", 
               random_answer_order = TRUE,
               button_label = "Check answer",
               right = "That's right, now only one of the four null hypotheses can be rejected.",
               wrong = "No, that's not it. Check the hint or the code below, if you're stuck.")
check_hint("You need to first extract the p-values from the outputs of the four *t*-tests and then apply the correction using the `p.adjust()` function.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

```{r}
#| code-fold: true
#| code-summary: "Show sample code to answer Q11.17"
#| eval: false

# Running the four t-tests and saving the output as four R objects to the local environment:
active.ttest <- t.test(formula = Active ~ Gender, 
       data = Dabrowska.data)

passive.ttest <- t.test(formula = Passive ~ Gender, 
       data = Dabrowska.data)

postmod.ttest <- t.test(formula = Postmod ~ Gender, 
       data = Dabrowska.data)

subjrel.ttest <- t.test(formula = SubRel ~ Gender, 
       data = Dabrowska.data)

# Extracting just the p-values from the test outputs
p.values <- c(active.ttest$p.value, passive.ttest$p.value, postmod.ttest$p.value, subjrel.ttest$p.value)

# The original p-values:
p.values

# Applying Holm's correction to the p-values
p_values.adjusted <- p.adjust(p.values, method = "holm")

# The adjusted p-values in scientific notation:
p_values.adjusted

# Now only the first p-value is below 0.05:
p_values.adjusted < 0.05
```
:::
### Check your progress 🌟 {.unnumbered}

Well done! You have successfully completed this chapter introducing the complex topic of inferential statistics. You have answered [`r checkdown::insert_score()` out of 17 questions]{style="color:green;"} correctly.
