# Ch. 7: Tasks & Quizzes {.unnumbered}

::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q7.1**]{style="color:green;"} The `View()` function is more user-friendly than attempting to examine the full table in the Console. Try to display the full L2.dataset in the Console by using the command `L2.data` which is shorthand for `print(L2.data)`. What happens?

```{r}
#| echo: false
#| label: "Q7.1"
library(checkdown)
check_question("R only displays the first 22 rows and the columns are not aligned because the Console window is not wide enough.", 
               options = c("R only displays the first 22 rows and the columns are not aligned because the Console window is not wide enough.", 
               "R produces an error message because there are too many rows and the Console window is not long enough.", 
               "The R Console prints the data in a randomly jumbled way.",
               "The R Console prints out all the data, but not the column headers."), 
               type = "radio",
button_label = "Check answer",
q_id = "Q7.1", 
random_answer_order = TRUE,
right = "That's right. The printed message \"[ reached 'max' / getOption(\"max.print\") -- omitted 45 rows ]\" is not an error, but merely a message informing us that not all the rows could be printed.",
wrong = "No, that's not what is happening. Try scrolling up within the Console window to better understand what is going on. Can you find the column headers? Check the hint below if you're stuck.")
check_hint("What happens when you increase or decrease the width of the Console pane in *RStudio* and then try to print the table again?", 
           hint_title = "🐭 Click on the mouse for a hint.")
```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q7.2**]{style="color:green;"} Six is the default number of rows printed by the `head()` function. Have a look at the function's help file using the command `?head` to find out how to change this default setting. How would you get `R` to print the first 10 lines of `L2.data`?

```{r}
#| echo: false
#| label: "Q7.2"

check_question(c("head(L2.data, n = 10)", "head(L2.data, n = 10L)", "head(L2.data, 10)"),
               options = c("head(L2.data, n = 10)", 
                           "head(L2.data n = 10L)",
                           "head(L2.data, n = 10L)",
                           "head(L2.data, L = 10)",
                           "head(L2.data, rows = 10)",
                           "head(L2.data, 10)"), 
               type = "radio",
button_label = "Check answer",
q_id = "Q7.2", 
random_answer_order = TRUE,
right = "That's right! In fact, three of the above options will work. Try them out to work out which three they are!",
wrong = "Humm, have you tried whether this really does display the first 10 rows of the dataset? Is the name of the argument that controls the number of rows to be printed correct? Are the arguments within the command separated by a comma?")
check_hint("You can find the correct command argument in the help file of the `head()` function. Remember that you can access a function's help file using the commands `?` or `help`.", hint_title = "🐭 Click on the mouse for a hint.")
```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q7.3**]{style="color:green;"} Which type of variable is stored in the `Occupation` column in `L1.data`?

```{r}
#| echo: false
#| label: "Q7.3"

check_question("Nominal",
               options = c("Nominal", 
                           "Binary",
                           "Continuous",
                           "Discrete",
                           "Other"),
button_label = "Check answer",
q_id = "Q7.3", 
random_answer_order = TRUE,
type = "radio",
right = "Correct. These job categories cannot be meaningfully ordered.",
wrong = "No, check the definitions above.")
```

[**Q7.4**]{style="color:green;"} Which type of variable is stored in the `Gender` column in `L1.data`?

```{r}
#| echo: false
#| label: "Q7.4"

check_question("Binary",
               options = c("Nominal", 
                           "Binary",
                           "Continuous",
                           "Discrete",
                           "Other"),
button_label = "Check answer",
q_id = "Q7.4", 
random_answer_order = TRUE,
type = "radio",
right = "That's right. Even though we know that human gender is not actually binary, in this dataset, the variable Gender is binary because all of the participants either identified as female (F) or male (M).",
wrong = "No, check the definitions above.")
```

[**Q7.5**]{style="color:green;"} Which type of variable is stored in the column `VocabR` in `L1.data`?

```{r}
#| echo: false
#| label: "Q7.5"

check_question("Discrete",
               options = c("Nominal", 
                           "Binary",
                           "Continuous",
                           "Discrete",
                           "Other"),
button_label = "Check answer",
q_id = "Q7.5", 
random_answer_order = TRUE,
type = "radio",
right = "That's right. These are participants' score on a vocabularly test. The more items they got right, the more points they got.",
wrong = "No, re-read the definitions above.")
```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q7.6**]{style="color:green;"} Use the `str()` function to examine the internal structure of the L2 dataset. How many columns are there in the L2 dataset?

```{r}
#| echo: false
#| label: "Q7.6"

check_question("45",
button_label = "Check answer",
q_id = "Q7.6", 
random_answer_order = TRUE,
right = "That's right, well done!",
wrong = "No, check the first line of the output of `str(L2.data)`.")
```

[**Q7.7**]{style="color:green;"} Which of these columns can be found in the L2 dataset, but not the L1 one?

```{r}
#| echo: false
#| label: "Q7.7"

check_question(c("NativeLg", "FirstExp", "Arrival", "EngWork"),
               options = c("NativeLg", "FirstExp", "Arrival", "EngWork", "EduYrs", "OtherLgs"), 
               type = "check",
button_label = "Check answer",
q_id = "Q7.7", 
random_answer_order = TRUE,
right = "That's right! And there are more columns that are specific to the L2 dataset but which are not listed here.",
wrong = "Not quite. Compare the outputs of `names(L1.data)` and `names(L2.data)` again.")
check_hint("Try using the `str()` or the `names()` functions to compare the two datasets.", hint_title = "🐭 Click on the mouse for a hint.")

```

[**Q7.8**]{style="color:green;"} Which type of `R` object is the variable `Arrival` stored as?

```{r}
#| echo: false
#| label: "Q7.8"

check_question("integer",
               options = c("integer",
                           "string character",
                           "digit",
                           "index",
                           "interest",
                           "intelligence"), 
               type = "check",
button_label = "Check answer",
q_id = "Q7.8", 
random_answer_order = TRUE,
right = "That's right! `int` is a subset of the numeric data type. It stands for 'integer' which is a whole number without a decimal place.",
wrong = "No, that's not it.")
check_hint("You can find the answer in the output of the command `str(L2.data)`", hint_title = "🐭 Click on the mouse for a hint.")

```

[**Q7.9**]{style="color:green;"} How old was the third participant listed in the L2 dataset when they first moved to an English-speaking country?

```{r}
#| echo: false
#| label: "Q7.9"

check_question("19", 
button_label = "Check answer",
q_id = "Q7.9", 
right = "Yes, well done!",
wrong = "No. Have a look at the third number printed after `$ Arrival` in the output of the command `str(L2.data)`")
check_hint("You can find the answer in the output of the command `str(L2.data)`", hint_title = "🐭 Click on the mouse for a hint.")
```

[**Q7.10**]{style="color:green;"} In both datasets, the column `Participant` contains anonymised participant IDs. Why is the variable `Participant` stored as string character vector in `L1.data`, but as an integer vector in `L2.data`?

```{r}
#| echo: false
#| label: "Q7.10"

check_question("Because some of the L1 participants' IDs contain letters as well as numbers.",
               options = c("Because some of the L1 participants' IDs contain letters as well as numbers.",
                           "Because there are more L1 participants than L2 participants.",
                           "Because the L1 participants' IDs only contain whole numbers with no decimal points.",
                           "Because the L1 participants' IDs are written out as words rather than digits."), 
               type = "check",
button_label = "Check answer",
q_id = "Q7.10", 
random_answer_order = TRUE,
right = "That's right! 🎉",
wrong = "No, that's not it. Have you had a look at the full set of participant IDs in the L1 dataset?")
check_hint("Print the full list of values in both columns using the commands `L1.data$Participant` and `L2.data$Participant` to work it out.", 
           hint_title = "🐭 Click on the mouse for a hint.")

```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

The following two quiz questions focus on the `NativeLg` variables from the L2 dataset (`L2.data`).

[**Q7.11**]{style="color:green;"} Use the index operators to find out the native language of the 26^th^ L2 participant.

```{r}
#| echo: false
#| label: "Q7.11"

check_question("Polish",
               options = c("Cantonese", "Chinese", "German", "Italian", "Greek", "Lithuanian", "Mandarin", "Polish", "Russian", "Spanish"),
button_label = "Check answer",
q_id = "Q7.11", 
right = "Brawo! 🇵🇱",
wrong = "No, that's not it. Check the hint if you need some help.")
check_hint("Use the command `L2.data$NativeLg[26]` to find out which it is.", 
           hint_title = "🐭 Click on the mouse for a hint.")
```

[**Q7.12**]{style="color:green;"} Which command(s) can you use to display only the Gender, Occupation, Native language, and Age of the last participant listed in the L2 dataset?

```{r}
#| echo: false
#| label: "Q7.12"

check_question(c("L2.data[67,c(2,3,5,9)]", 
                 "L2.data[67,c(2:3,5,9)]", 
                 "L2.data[67 , c(2,3,5,9)]"),
               options = c("L2.data[67,c(2,3,5,9)]", "L2.data[67 , c(2,3,5,9)]", "L2.data[67,c(2:3,5,9)]", "L2.data[67:c(2,3,5,9)]", "L2.data[-1:c(2,3,5,9)]", "L2.data[90,c(2:3,5,9)]"),
button_label = "Check answer",
q_id = "Q7.12", 
type = "check",
random_answer_order = TRUE,
right = "Great job!",
wrong = "Not quite. Try out these options to find out for yourself which ones work and which don't.")
check_hint("Using `-1` as an index to access the last row is a neat trick in Python, but it sadly doesn't work in `R`.", hint_title = "🦉 Hover over the owl for a hint", type = "onmouseover")
check_hint("Three of these will work. You'll need to find all three!", 
           hint_title = "🐭 Click on the mouse for a second hint.")

```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q7.13**]{style="color:green;"} How does the average age of the L2 participants in @DabrowskaExperienceAptitudeIndividual2019 compare to that of the L1 participants?

```{r}
#| echo: false
#| label: "Q7.13"

check_question("On average, the L2 participants are younger than the L1 participants.", 
               options = c("On average, the L2 participants are younger than the L1 participants.", 
               "On average, the L2 participants are older than the L1 participants.", 
               "On average, the L2 participants are the same age than the L1 participants.",
               "Age is not comparable across two different datasets."), 
               type = "radio",
button_label = "Check answer",
q_id = "Q7.13", 
random_answer_order = TRUE,
right = "That's right!",
wrong = "No, this is incorrect. Compare the results of the commands `mean(L2.data$Age)` and `mean(L1.data$Age)`.")
```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

For this [**task**]{style="color:green;"}, you first need to check that you have saved the following two variables from the L1 dataset to your `R` environment.

```{r}
#| eval: false
L1.Age <- L1.data$Age
L1.Occupation <- L1.data$Occupation
```

[**Q7.14**]{style="color:green;"} Below is a list of useful base `R` functions. Try them out with the variable `L1.Age`. What does each function do? Make a note by writing a comment next to each command (see @sec-Comments). The first one has been done for you.

```{r eval = FALSE}
mean(L1.Age) # The mean() function returns the mean average of a set of number.
min()
max()
sort()
length()
mode()
class()
table()
summary()

```

[**Q7.15**]{style="color:green;"} `Age` is a numeric variable. What happens if you try these same functions with a character string variable? Find out by trying them out with the variable `L1.Occupation` which contains words rather than numbers.

:::
::: {.callout-note collapse="true"}
#### Click here for the solutions to [**Q7.14**]{style="color:green;"}—[**Q7.15**]{style="color:green;"}.

As you will have seen, often the clue is in the name of the function - but not always! 😉

Hover your mouse over the numbers on the right for the solutions to appear.

```{r}
#| eval: false

mean(L1.Age) # <1>
mean(L1.Occupation) # <2>

min(L1.Age) # <3>
min(L1.Occupation) # <4>

max(L1.Age) # <5>
max(L1.Occupation) # <6>

sort(L1.Age) # <7>
sort(L1.Occupation) # <8>

length(L1.Age) # <9>
length(L1.Occupation) # <10>

mode(L1.Age) # <11>
mode(L1.Occupation) # <12>

class(L1.Age) # <13>
class(L1.Occupation) # <14>

table(L1.Age) # <15>
table(L1.Occupation) # <16>

summary(L1.Age) # <17>
summary(L1.Occupation) # <18>
```

1.  The `mean()` function returns the mean average of a set of number.
2.  It does not make sense to calculate a mean average value of a set of words, therefore `R` returns an `NA` (not applicable) and a warning in red explaining that the `mean()` function expects a numeric or logical argument.
3.  For a numeric variable, `min()` returns the lowest numeric value.
4.  For a string variable, `min()` returns the first value sorted alphabetically.
5.  For a numeric variable, `min()` returns the highest numeric value.
6.  For a string variable, `max()` returns the last value sorted alphabetically.
7.  For a numeric variable, `sort()` returns all of the values of the variable ordered from the smallest to the largest.
8.  For a string variable, `sort()` returns of all of the values of the variable in alphabetical order.
9.  The function `length()` returns the number of values in the variable.
10. The function `length()` returns the number of values in the variable.
11. The function `mode()` returns the `R` data type that the variable is stored as.
12. The function `mode()` returns the `R` data type that the variable is stored as.
13. The function `mode()` returns the `R` object class that the variable is stored as.
14. The function `mode()` returns the `R` object class that the variable is stored as.
15. For a numeric variable, the function `table()` outputs a table that tallies the number of occurrences of each unique value in a set of values and sorts them in ascending order.
16. For a string variable, the function `table()` outputs a table that tallies the number of occurrences of each unique value in a set of values and sorts them alphabetically.
17. For a numeric variable, the function `summary()` outputs six values that, together, summarise the set of values contained in this variable: the minimum and maximum values, the first and third quartiles, and the mean and median (see Chapter 8).
18. For a string variable, the `summary()` function only outputs the length of the string vector, its object class and data mode.
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

Look at the following two lines of code and their (abbreviated) outputs.

```{r include=FALSE}
library(here)
library(tidyverse)

L1.data <- read.csv(file = here("data", "L1_data.csv"))
L2.data <- read.csv(file = here("data", "L2_data.csv"))
```

```{r}
#| eval: false
L1.data$Vocab
```

```{r}
#| echo: false
head(L1.data$Vocab)
```

```{r}
#| eval: false
round(L1.data$Vocab)
```

```{r}
#| echo: false
head(round(L1.data$Vocab))
```

[**Q7.16**]{style="color:green;"} Based on your observations, what does the `round()` function do?

```{r}
#| echo: false
#| label: "Q7.16"

check_question("The round() function rounds off numbers to the nearest whole number.", 
               options = c("The round() function displays just the first two digits of any number.", 
                           "The round() function rounds off numbers to the nearest whole number.",
               "The round() function is designed to save screen space for smaller displays.", 
               "The round() function displays fewer values for ease of reading."),
               type = "radio",
button_label = "Check answer",
q_id = "Q7.16", 
random_answer_order = TRUE,
right = "That's right.",
wrong = "Are you sure? Try out the `round()` function with other numbers to find out more.")
```

[**Q7.17**]{style="color:green;"} Check out the 'Usage' section of the help file on the `round()` function to find out how to round the `Vocab` values in the L1 dataset to two decimal places. How can this be achieved?

```{r}
#| echo: false
#| label: "Q7.17"
check_question(c("round(L1.data$Vocab, 2)",
                 "round(L1.data$Vocab, digits = 2)"),
               options = c("round(L1.data$Vocab, 2)",
                           "round(L1.data$Vocab, digits = 2)",
                           "round(L1.data$Vocab: digits = 2)",
                           "round(L1.data$Vocab, digits = -4)"), 
               type = "check",
button_label = "Check answer",
q_id = "Q7.17", 
random_answer_order = TRUE,
right = "That's right.",
wrong = "Not quite. Try out these commands to find out which ones work as expected.")
check_hint("Two options are correct because, as long as the arguments are listed in the correct order, we do not have to specify the names of the arguments.", 
           hint_title = "🐭 Click on the mouse for a hint.")

```
:::
::: {.callout-tip collapse="false"}
#### Your turn! {.unnumbered}

[**Q7.18**]{style="color:green;"} Using the `R` pipe operator, calculate the average mean age of the L2 participants and round off this value to two decimal places. What is the result?

```{r}
#| echo: false
#| label: "Q7.18"

check_question("32.72", 
button_label = "Check answer",
random_answer_order = TRUE,
right = "Yes, well done!",
wrong = "No, try again.")
```

[**Q7.19**]{style="color:green;"} Unsurprisingly, in @DabrowskaExperienceAptitudeIndividual2019's study, English L1 participants, on average, scored higher in the English vocabulary test than L2 participants. Calculate the difference between L1 and L2 participants' mean `Vocab` test results and round off this means difference to two decimal places.

```{r}
#| echo: false
#| label: "Q7.19"

check_question("16.33",
button_label = "Check answer",
q_id = "Q7.19", 
random_answer_order = TRUE,
right = "That's right. Very well done!",
wrong = "That's not it. Check the code chunk below if you are unsure how to proceed.")
check_hint("There are several ways to solve this puzzle. It does not have to involve the pipe operator. Don't hesitate to work in several steps by saving an intermediary result as an `R` object.", hint_title = "🐭 Click on the mouse for a hint.")
```
:::
::: {.callout-note collapse="true"}
#### Click here for a detailed answer to [**Q7.19**]{style="color:green;"}

They are lots of ways to tackle this in `R`. Here is a first approach that involves the pipe operator:

```{r include=FALSE}
library(here)
library(tidyverse)

L1.data <- read.csv(file = here("data", "L1_data.csv"))
L2.data <- read.csv(file = here("data", "L2_data.csv"))
```

```{r}
(mean(L1.data$Vocab) - mean(L2.data$Vocab)) |> 
  round(digits = 2)
```

Note that this approach requires a set of brackets around the first subtraction operation, otherwise only the second mean value is rounded off to two decimal places. Compare the following lines of code:

```{r}

mean(L1.data$Vocab) - mean(L2.data$Vocab)

(mean(L1.data$Vocab) - mean(L2.data$Vocab)) |> 
  round(digits = 2)

mean(L1.data$Vocab) - round(mean(L2.data$Vocab), digits = 2)

```

An alternative approach would be to store the difference in means as an `R` object and, in a second line of code, pass this object to the `round()` function.

```{r}
mean.diff.vocab <- mean(L1.data$Vocab) - mean(L2.data$Vocab)

round(mean.diff.vocab, digits = 2)
```

Or, you could combine both approaches like this:

```{r}
mean.diff.vocab <- mean(L1.data$Vocab) - mean(L2.data$Vocab)
mean.diff.vocab |> 
  round(digits = 2)

```

There is often more than one way to solve problems in `R`. Choose whichever way you are most comfortable with. As you long as you understand what your code does (see @sec-AI), it doesn't matter if it's particularly elegant or efficient.
:::
### Check your progress 🌟 {.unnumbered}

You have successfully completed [`r checkdown::insert_score()` out of 19 questions]{style="color:green;"} in this chapter.
